A Formal Verification Framework for Tezos Smart Contracts Based on Symbolic Execution

### 2024-08-16 papers/static-checker/aplas/paper.tex
camera-ready

% APLAS 2021: Regular research papers should not exceed 18 pages in
% the Springer LNCS format(LaTeX template), including bibliography and
% figures.
% Lightweight double-blind: Author names and institutions must be
% omitted and References to the authors’ own related work should be in
% the third person 
%
\documentclass[runningheads]{llncs}
\pdfoutput=1
%\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{comment}
\usepackage{graphicx}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{mathpartir}
%\usepackage{fixme}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{adjustbox} % For adjusting box sizes
\lstdefinelanguage{michelson}{
  % basicstyle=\ttfamily,
  basicstyle=\fontsize{8}{9.6}\selectfont,
  morekeywords={parameter,storage,or,unit,mutez,pair,bool,address}, sensitive=false,
  morecomment=[l]{\#},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
}
\lstset{
  language=Caml,
  captionpos=b,
  aboveskip=-\smallskipamount,
  belowskip=-\smallskipamount,
  belowcaptionskip=0pt,
  basicstyle=\fontsize{8}{9.6}\selectfont,
  morekeywords={val}
}

@@

\begin{document}
%
\title{A Formal Verification Framework for Tezos Smart Contracts Based on Symbolic Execution}
%
\titlerunning{A Formal Verification Framework for Tezos Smart Contracts}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Thi Thu Ha Doan\orcidID{0000-0001-7524-4497}\and Peter Thiemann\orcidID{0000-0002-9000-1239}}

%
%\authorrunning{Ha Doan, P. Thiemann}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{University of Freiburg, Germany \\
%  \email{\{doanha,thiemann\}@informatik.uni-freiburg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We developed SCV, a verification tool geared towards contracts implemented in Michelson, the smart contract language of the Tezos blockchain. SCV utilizes symbolic execution to derive verification conditions that can be checked automatically against user specifications using an SMT solver. These specifications are written in a domain-specific language. We present the fundamental principles behind the design and development of SCV. By pruning states during symbolic execution, SCV effectively addresses the problem of state explosion. This approach has enabled us to conduct two case studies on real-life contracts, demonstrating the expressiveness of our tool.
\keywords{Smart Contract, Blockchain, Formal Verification, Symbolic Execution}
\end{abstract}
\section{Introduction}
\label{sec:introduction}
Since smart contracts are permanently stored on the blockchain, once deployed, a smart contract cannot be changed. Regrettably, numerous instances have been documented where vulnerabilities in smart contracts were identified by the blockchain community, resulting in substantial financial losses~\cite{dao,wallethack}. The distinctive execution model of smart contract languages has given rise to unforeseen errors due to a lack of familiarity with their complexities. These factors, combined with the immutability of contracts, underscore the need to ensure their correctness prior to deployment.

@@

Several high-level languages, such as LIGO \cite{ligo}, SmartPy \cite{smartpy}, and Liquidity \cite{liquidity}, provide a more readable syntax resembling popular programming languages. This allows developers to write Tezos smart contracts more comfortably. Our tools are applicable to all these languages because they compile to Michelson.
\section{Michelson}
\label{sec:background}
\lstset{language=michelson}
\begin{lstlisting}[float=tp,captionpos=b,caption={A Michelson program},label={lst:simple-program-add}]
parameter int;
storage   int;
code {                (* pair int int :: [] *)
  DUP; UNPAIR;        (* int :: int :: pair int int :: [] *) 
  COMPARE; GE;        (* bool :: pair int int :: [] *)
  IF {UNPAIR; SUB} {PUSH string 'Unexpected Pair'; FAILWITH};
                      (* int :: [] *)
  NIL operation; PAIR (* pair (list operation) int :: [] *)
}
\end{lstlisting}
Michelson is a stack-based, statically typed functional programming language. A program is defined by the type $S$ of its storage, the type $P$ of its single parameter, and a typed list of instructions for the underlying stack machine. It starts on a stack with one element containing a pair of the current storage value (taken from the blockchain) and the parameter value specified by the contract invocation. When the program finishes, the stack must contain a single element with a pair of the final storage value (to be saved on the blockchain) and a list of operations. These operations are syntactic commands, such as token transfers or contract invocations, among other possibilities. Thus, the program computes a function $P \times S \to O^* \times S$, where $O$ is the type of operations.

Besides the standard datatypes (e.g., numbers, lists, pairs, sums, functions) with the expected operations, Michelson supports a number of types and operations specific to blockchain applications (e.g., \texttt{mutez} for tokens) as shown in Fig.~\ref{fig:types}, where \textbf{cpt} denotes comparable types. The type system does not support quantification over types, and there are no user-defined types.

@@

A contract may have several entrypoints. Entrypoints are encoded by
using a sum type (call ``or type'' in Michelson). Each (nested) alternative in a sum can be
given a name, which is used as the name for an entrypoint. A call to
an entrypoint automatically adds the required \lstinline|LEFT| and
\lstinline|RIGHT| constructors to the parameter.
\lstset{language=Caml}
\section{SCV: A Domain Specific Language For Smart Contract Property Specification}
\label{sec:domain-specific-language}
SCV contains a domain-specific language to specify smart contract properties, aiming to be user-friendly for non-experts and compatible with SMT solvers. 
\begin{figure}[tp]
\begin{align*}
\textbf{cpt} &::= 
   \text{unit} 
   \Mid\text{never} 
   \Mid\text{bool} 
   \Mid\ \text{int}
   \Mid\text{nat}
   \Mid\text{string}
   \Mid\text{chain-id} 
    \Mid\text{bytes}
   \Mid\text{mutez} 
   \Mid\text{key-hash} \\
   &\Mid\text{key}
   \Mid \text{signature}
   \Mid \text{timestamp} 
   \Mid \text{address} 
   \Mid \text{option}\ \textbf{cpt}
   \Mid \text{or}\ \textbf{cpt}\ \textbf{cpt}
   \Mid \text{pair}\ \textbf{cpt}\ \textbf{cpt} \\
\text{T, U} &::= 
    \text{cpt}
   \Mid \text{option}\ \text{T}
   \Mid \text{list}\ \text{T}
   \Mid \text{set}\ \textbf{cpt} 
   \Mid \text{operation} 
   \Mid \text{contract}\ \text{T} \\
  & \Mid \text{pair}\ \text{T}\ \text{U}
   \Mid \text{or}\ \text{T}\ \text{U}
   \Mid \text{lambda}\ \text{T}\ \text{U} 
   \Mid \text{map}\ \textbf{cpt}\ \text{T}
   \Mid \text{big-map}\ \textbf{cpt}\ \text{T}
\end{align*}
\caption{Types}
\label{fig:types}
\end{figure}
\subsection{Basic syntax}
\label{sec:basic-syntax}
Listing~\ref{lst:sub-contract-specification} shows the specification of the \texttt{sub} contract in the SCV language. Here, \KMCONTRACT, \KSPEC, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are keywords. The \KINPUT{} and \KOUTPUT{} fields specify the input and output patterns, which are terms on Michelson values.

@@

Since Michelson operates on a stack, we assume that the input stack and the output stack each have only one element, as specified in the input term and the output term, respectively. The syntax allows us to specify both incomplete and complete smart contracts. For a complete smart contract, the output should be a pair consisting of an operation list and the storage's state after the completion of the code. For verification purposes, we elide the operation list and represent the output solely by the term of the value in the storage after execution concludes. %Michelson code, whether composed by users, obtained from the Tezos blockchain, or compiled from higher-level languages, can be seamlessly used within our framework.

The pre- and post-condition instances are formulas in first-order logic that can be handled by an SMT solver (specifically, Z3 in our case). The specification language is constructed from predicates. Formulas used in assertions or added to solvers are terms of Boolean type. Non-Boolean terms may also be used, and they can be mixed in any combination where types match appropriately.

@@

Certain functions are defined as predicates, such as \text{transfer\_token(e1, e2, e3)}. This predicate holds true if, after executing the smart contract code, there is an operation in the operation list indicating the transfer of \text{e2} tokens to address \text{e3} along with parameter \text{e1}\footnote{Additional predicates can be straightforwardly added if a mapping to Z3 formulas is defined.}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of  the \lstinline/sub/ contract},label={lst:sub-contract-specification},numbers=left]
mcontract sub = spec
  code := {DUP; UNPAIR; COMPARE; GE; IF {UNPAIR; SUB} 
          {PUSH string 'Unexpected Pair'; FAILWITH}}
  input := Pair (x : int) (y : int)
  output := (z : int)
  pre-condition := (x >= y)
  post-condition := (z = x - y) && (z >= 0)
\end{lstlisting}

@@

\lstinline/z/ is equal to the substract of

@@

at once and to describe relationships between entrypoints. The entrypoint name is written after the \lstinline/%/ symbol and is followed by a property specification for the entrypoint.  The specification finishes with specifying entrypoint relations, which we explain next. For a smart contract, the input term includes the storage term before, and the output term represents the storage after the execution. Thus, there are two options of the syntax. The first option similar to the basic syntax with the entrypoint specific.  

@@

Let us have a look at the specification for the auction
contract that models an online auction in Listing \ref{lst:auction-contract-specification1}. The contract has two entry points, \lstinline|%bid| and \lstinline|%close|, which serve as functions for bidding and closing the contract, respectively. The smart contract storage is a pair containing, as its first element, a boolean value \lstinline|open| that indicates whether the contract is still open for bidding or already closed. The second element is another nested pair, where the first element \lstinline|highest| stores the highest bidder's address, and the second element \lstinline|owner| is the contract owner's address. The current highest bid is reflected in the balance of the contract before the next bid is invoked. We omit the Michelson code of the smart contract due to its length.  For the \lstinline/%bid/ entrypoint, the parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is described by the \lstinline|storage| phrase.

Closing the contract transfers the balance to the owner and is restricted to the owner. Both closing and bidding operations fail if the auction is already closed. If bidding is open and the amount of tokens accompanying the bid exceeds the current highest bid, the current bidder replaces the previous highest bidder, and the previous highest bidder is reimbursed. Otherwise, bidding fails. Upon deployment, the owner deposits an initial balance to indicate the minimum bid. The contract should be deployed with the storage value \lstinline|Pair True (Pair owner owner)|, indicating that bidding is allowed, and the contract owner is currently the highest bidder.

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification1},numbers=left]
mcontract auction = spec 
 storage := Pair (open: bool) 
                 (Pair (highest: address) (owner: address))
 entrypoint %bid
  code := {...}
  parameter := Unit 
  pre-condition := open = true && Amount > pre_Balance
   && get_contract (unit, highest) = Some (c: contract unit)
  post-condition := open = true && post_highest = Sender 
   && post_Balance = Amount && post_owner = owner 
   && transfer_token (Unit, pre_Balance, c);
 entrypoint %close
  ...
 (%create -> %bid) with ...
 | (%create  -> %close) with ...
 | (%bid -> %bid) with (Amount > pre_Balance) 
  && get_contract (unit, highest) = Some (c: contract unit)
 | (%bid -> %close) with (Sender = owner) 
  && get_contract (unit, owner) = Some (c: contract unit)
 | not (%close -> %bid) | not (%close -> %close)
\end{lstlisting}

%There are several properties that we want to guarantee for this contract. 
The first property is that if a user calls the \lstinline|%bid| entrypoint when the contract is still open for bidding and the bid is higher than the current bid (as specified in the \lstinline|pre-condition| field at line 7), and the address of the highest bidder can be cast to a real contract represented by the variable \lstinline|c| of type \lstinline|contract unit| on the blockchain (line 8), then the bid should succeed. In this case, the highest bidder must be updated to the sender, which is the address of the caller (line 9 in the \lstinline|post-condition| field), and the balance of the contract should be the amount sent along with the call (line 10). Moreover, the previous highest bidder should receive their bid back, which means the auction contract should transfer the \lstinline|pre_Balance| to the previous bidder, \lstinline|highest| (line 11). Another property guarantees that the auction owner's address should not be changed after calling the \lstinline/bid/ entrypoint (line 10) and the auction should be still open (line 9). 
\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}
Since the order of smart contract
calls can be crucial, we specify ordering constraints among calls with
 entrypoint relations. An entrypoint relation can be specified as
\lstinline/%a -> %b with p/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that from the entrypoint
that satisfies the condition \lstinline/p/, it is possible  to call
the entrypoint \lstinline/%b/. One specification can
contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

Let us return to the auction contract. It is important to guarantee
that  the bid call can happen successfully after another if the new bid is higher than the previous
one (lines 16-17). The  \lstinline/%close/ entrypoint can be called
anytime after bidding if the caller is the contract owner (lines 18-19). However, it
should not be possible to call the  \lstinline/%bid/ entrypoint after
the auction is closed, and, of course, it cannot be closed
again if it is already closed (line 20).  Furthermore, contract deployment sets the  \lstinline/open/ flag to
\TRUE, indicating that the contract is open for bidding, i.e., users
should be able to bid on the contract or the contract owner should be
able to close it. Otherwise, the owner's original
deposit is lost. This sequence  illustrates
the life cycle of the auction contract, which is depicted in Figure
\ref{fig:auction-life-cycle}. Other properties related to the \lstinline|%close| entrypoint and the life cycle are omitted here due to page limits. %, but the full specification can be found in the Appendix and is available online at \cite{scv}.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.9\textwidth]{auction}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

\section{Static Checker}
\label{sec:static-checker-we}
The architecture of the static checker is illustrated in Figure~\ref{fig:architecture-of-static-checker}. Users write specifications in the SCV language, which are then parsed into an Abstract Syntax Tree (AST) in OCaml. The input term along with the code is passed to the symbolic interpreter, while the output term and the pre and postconditions are fed to the static checker. The symbolic interpreter, which is fully typed and rejects ill-typed programs, then runs symbolic execution on the code with the provided input, where the initial stack is derived from the input term.

>>>>

% APLAS 2024: Regular research papers should not exceed 17 pages in
% the Springer LNCS format(LaTeX template), including bibliography and
% figures.
% Lightweight double-blind: Author names and institutions must be
% omitted and References to the authors’ own related work should be in
% the third person 
%
\documentclass[runningheads]{llncs}
\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathpartir}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}

\lstdefinelanguage{michelson}{
  basicstyle=\ttfamily,
  basicstyle=\fontsize{8}{9.6}\selectfont,
  morekeywords={parameter,storage,or,unit,mutez,pair,bool,address}, sensitive=false,
  morecomment=[l]{\#},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
}

@@

\begin{document}
%
\title{A Formal Verification Framework for Tezos Smart Contracts Based on Symbolic Execution}
%
\titlerunning{A Formal Verification Framework for Tezos Smart Contracts}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Thi Thu Ha Doan\orcidID{0000-0001-7524-4497}\and Peter Thiemann\orcidID{0000-0002-9000-1239}}

%
\authorrunning{Ha Doan, P. Thiemann}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Freiburg, Germany \\
  \email{\{doanha,thiemann\}@informatik.uni-freiburg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We developed SCV, a verification tool geared towards contracts implemented in Michelson, the smart contract language of the Tezos blockchain. SCV utilizes symbolic execution to derive verification conditions that can be checked automatically against user specifications using an SMT solver. These specifications are written in a domain-specific language. We present the fundamental principles behind the design and development of SCV. By pruning states during symbolic execution, SCV effectively addresses the problem of state explosion. This approach has enabled us to conduct two case studies on real-life contracts, demonstrating the expressiveness of our tool.
\keywords{Smart Contract, Blockchain, Formal Verification, Symbolic Execution}
\end{abstract}
\section{Introduction}
\label{sec:introduction}
Since smart contracts are permanently stored on the blockchain, once deployed, a smart contract cannot be changed. Regrettably, numerous instances have been documented where vulnerabilities in smart contracts were identified by the blockchain community, resulting in substantial financial losses~\cite{dao,wallethack}. The distinctive execution model of smart contract languages has given rise to unforeseen errors due to a lack of familiarity with their complexities. These factors, combined with the immutability of contracts, underscore the need to ensure their correctness prior to deployment.

@@

Several high-level languages, such as LIGO \cite{ligo}, SmartPy \cite{smartpy}, and Liquidity \cite{liquidity}, provide a more readable syntax resembling popular programming languages. This allows developers to write Tezos smart contracts more comfortably. Our tools are applicable to all these languages because they compile to Michelson.
\section{Michelson}
\label{sec:background}
Michelson is a stack-based, statically typed functional programming language. A program is defined by the type $S$ of its storage, the type $P$ of its single parameter, and a typed list of instructions for the underlying stack machine. It starts on a stack with one element containing a pair of the current storage value (taken from the blockchain) and the parameter value specified by the contract invocation. When the program finishes, the stack must contain a single element with a pair of the final storage value (to be saved on the blockchain) and a list of operations. These operations are syntactic commands, such as token transfers or contract invocations, among other possibilities. Thus, the program computes a function $P \times S \to O^* \times S$, where $O$ is the type of operations.

Besides the standard datatypes (e.g., numbers, lists, pairs, sums, functions) with the expected operations, Michelson supports a number of types and operations specific to blockchain applications (e.g., \texttt{mutez} for tokens) as shown in Fig.~\ref{fig:types}, where \textbf{cpt} denotes comparable types. The type system does not support quantification over types, and there are no user-defined types.
\begin{figure}[tp]
\begin{align*}
\textbf{cpt} &::= 
   \text{unit} 
   \Mid\text{never} 
   \Mid\text{bool} 
   \Mid\ \text{int}
   \Mid\text{nat}
   \Mid\text{string}
   \Mid\text{chain-id} 
    \Mid\text{bytes}
   \Mid\text{mutez} 
   \Mid\text{key-hash} \\
   &\Mid\text{key}
   \Mid \text{signature}
   \Mid \text{timestamp} 
   \Mid \text{address} 
   \Mid \text{option}\ \textbf{cpt}
   \Mid \text{or}\ \textbf{cpt}\ \textbf{cpt}
   \Mid \text{pair}\ \textbf{cpt}\ \textbf{cpt} \\
\text{T, U} &::= 
    \text{cpt}
   \Mid \text{option}\ \text{T}
   \Mid \text{list}\ \text{T}
   \Mid \text{set}\ \textbf{cpt} 
   \Mid \text{operation} 
   \Mid \text{contract}\ \text{T} \\
  & \Mid \text{pair}\ \text{T}\ \text{U}
   \Mid \text{or}\ \text{T}\ \text{U}
   \Mid \text{lambda}\ \text{T}\ \text{U} 
   \Mid \text{map}\ \textbf{cpt}\ \text{T}
   \Mid \text{big-map}\ \textbf{cpt}\ \text{T}
\end{align*}
\caption{Types}
\label{fig:types}
\end{figure}

@@

A contract may have several entrypoints. Entrypoints are encoded by
using a sum type (call ``or type'' in Michelson). Each (nested) alternative in a sum can be
given a name, which is used as the name for an entrypoint. A call to
an entrypoint automatically adds the required \lstinline|LEFT| and
\lstinline|RIGHT| constructors to the parameter.
\lstset{language=michelson}
\begin{lstlisting}[float,captionpos=b,caption={A Michelson program},label={lst:simple-program-add}]
parameter int;
storage   int;
code {                (* pair int int :: [] *)
  DUP; UNPAIR;        (* int :: int :: pair int int :: [] *) 
  COMPARE; GE;        (* bool :: pair int int :: [] *)
  IF {UNPAIR; SUB} {PUSH string 'Unexpected Pair'; FAILWITH};
                      (* int :: [] *)
  NIL operation; PAIR (* pair (list operation) int :: [] *)}
\end{lstlisting}
\section{SCV: A Domain Specific Language For Smart Contract Property Specification}
\label{sec:domain-specific-language}
SCV contains a domain-specific language to specify smart contract properties, aiming to be user-friendly for non-experts and compatible with SMT solvers. 
\subsection{Basic syntax}
\label{sec:basic-syntax}
Listing~\ref{lst:sub-contract-specification} shows the specification of the \texttt{sub} contract in the SCV language. Here, \KMCONTRACT, \KSPEC, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are keywords. The \KINPUT{} and \KOUTPUT{} fields specify the input and output patterns, which are terms on Michelson values.

@@

Since Michelson operates on a stack, we assume that the input stack and the output stack each have only one element, as specified in the input term and the output term, respectively. The syntax allows us to specify either a fragment of Michelson code or a complete program. For a complete Michelson program, the output should be a pair consisting of an operation list and the state of the storage after code execution,  as described in Section \ref{sec:background}. For verification purposes, we elide the operation list and represent the output solely by the term of the value in the storage after execution concludes.

The pre and postcondition instances are formulas in first-order logic that can be handled by an SMT solver (specifically, Z3 in our case). The specification language is constructed from predicates. Formulas used in assertions or added to solvers are terms of Boolean type. Non-Boolean terms may also be used, and they can be mixed in any combination where types match appropriately.

@@

Certain functions are defined as predicates, such as \text{transfer\_token(e1, e2, e3)}. This predicate holds true if, after executing the smart contract code, there is an operation in the operation list indicating the transfer of \text{e2} tokens to address \text{e3} along with parameter \text{e1}\footnote{Additional predicates can be straightforwardly added if a mapping to Z3 formulas is defined.}.
%\lstset{language=michelson}
\begin{lstlisting}[float,captionpos=b,caption={Specification of  the \lstinline/sub/ contract},label={lst:sub-contract-specification},numbers=left]
mcontract sub = spec
  code := {DUP; UNPAIR; COMPARE; GE; IF {UNPAIR; SUB} 
          {PUSH string 'Unexpected Pair'; FAILWITH}}
  input := Pair (x : int) (y : int)
  output := (z : int)
  pre-condition := (x >= y)
  post-condition := (z = x - y) && (z >= 0)
\end{lstlisting}

@@

\lstinline/z/ is equal to the subtract of

@@

at once and to describe relationships between entrypoints. The entrypoint name is written after the \lstinline/%/ symbol and is followed by a property specification for the entrypoint.  The specification finishes with specifying entrypoint relations, which we explain next. For a smart contract, the input term includes the storage term before, and the output term represents the storage after the execution. Thus, there are two options of the syntax. The first option is similar to the basic syntax with the entrypoint specific.  

@@

Let us have a look at the specification for the auction
contract that models an online auction in Listing \ref{lst:auction-contract-specification1}. The contract has two entry points, \lstinline|%bid| and \lstinline|%close|, which serve as functions for bidding and closing the contract, respectively. The smart contract storage is a pair containing, as its first element, a boolean value \lstinline|open| that indicates whether the contract is still open for bidding or already closed. The second element is another nested pair, where the element \lstinline|highest| stores the highest bidder's address, and the element \lstinline|owner| is the contract owner's address. The current highest bid is reflected in the balance of the contract before the next bid is invoked. We omit the Michelson code of the smart contract due to its length.  For the \lstinline/%bid/ entrypoint, the parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is described by the \lstinline|storage| phrase.

Closing the contract transfers the balance to the owner and is restricted to the owner. Both closing and bidding operations must fail if the auction is already closed. If bidding is open and the amount of tokens accompanying the bid exceeds the current highest bid, the current bidder replaces the previous highest bidder, and the previous highest bidder is reimbursed. Otherwise, bidding fails. Upon deployment, the owner deposits an initial balance to indicate the minimum bid. The contract should be deployed with the storage value \lstinline|Pair True (Pair owner owner)|, indicating that bidding is allowed, and the contract owner is currently the highest bidder.
\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification1},numbers=left]
mcontract auction = spec 
 storage := Pair (open: bool) 
                 (Pair (highest: address) (owner: address))
 entrypoint %bid
  code := {...}
  parameter := Unit 
  pre-condition := open = true && Amount > pre_Balance
   && get_contract (unit, highest) = Some (c: contract unit)
  post-condition := open = true && post_highest = Sender 
   && post_Balance = Amount && post_owner = owner 
   && transfer_token (Unit, pre_Balance, c);
 entrypoint %close
  ...
 (%create -> %bid) with ...
 | (%create  -> %close) with ...
 | (%bid -> %bid) with (Amount > pre_Balance) 
  && get_contract (unit, highest) = Some (c: contract unit)
 | (%bid -> %close) with (Sender = owner) 
  && get_contract (unit, owner) = Some (c: contract unit)
 | not (%close -> %bid) | not (%close -> %close)
\end{lstlisting}

The first property is that if a user calls the \lstinline|%bid| entrypoint when the contract is still open for bidding and the bid is higher than the current bid (as specified in the \lstinline|pre-condition| field at line 7), and the address of the highest bidder can be cast to a real contract represented by the variable \lstinline|c| of type \lstinline|contract unit| on the blockchain (line 8), then the bid should succeed. In this case, the highest bidder must be updated to the sender, which is the address of the caller (line 9 in the \lstinline|post-condition| field), and the balance of the contract should be the amount sent along with the call (line 10). Moreover, the previous highest bidder should receive their bid back, which means the auction contract should transfer the \lstinline|pre_Balance| to the previous bidder, \lstinline|highest| (line 11). Another property guarantees that the auction owner's address should not be changed after calling the \lstinline/bid/ entrypoint (line 10) and the auction should be still open (line 9). 
\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}
Since the order of smart contract
calls can be crucial, we specify ordering constraints among calls with
 entrypoint relations. An entrypoint relation can be specified as
\lstinline/%a -> %b with p/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that after invoking the entrypoint \lstinline/%a/, it is possible  to call
the entrypoint \lstinline/%b/ if the condition \lstinline/p/ holds. One specification can contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

Let us return to the auction contract. It is important to guarantee
that  the bid call can happen successfully after another if the new bid is higher than the previous
one (lines 16-17). The  \lstinline/%close/ entrypoint can be called
anytime after bidding if the caller is the contract owner (lines 18-19). However, it
should not be possible to call the  \lstinline/%bid/ entrypoint after
the auction is closed, and, of course, it cannot be closed
again if it is already closed (line 20).  Furthermore, contract deployment sets the  \lstinline/open/ flag to
\TRUE, indicating that the contract is open for bidding, i.e., users
should be able to bid on the contract or the contract owner should be
able to close it. Otherwise, the owner's original
deposit is lost. This sequence  illustrates
the life cycle of the auction contract, which is depicted in Figure
\ref{fig:auction-life-cycle}. Other properties related to the \lstinline|%close| entrypoint and the life cycle are omitted here due to page limits.
\begin{figure}[tp]
    \centering
    \includegraphics[width=0.9\textwidth]{auction}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model} 
We already learned about Michelson, the low-level smart contract
language for the Tezos blockchain, and its stack-based nature in
Section~\ref{sec:background}.  Here we address the changes needed for symbolic execution of Michelson programs, focusing on a selection of instructions. 
\subsection{System Model}
\label{sec:system-model}
Our symbolic model of a program state consists of a stack where each element is a symbolic value, represented as a pair comprising a Michelson term and its type. Thus, we write a stack as \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK. Here, \ensuremath{[\ ]} represents the empty stack, and \ensuremath{hd :: tl} denotes a stack with \ensuremath{hd} as the top element and \ensuremath{tl} as the remaining stack. Let \INSTRUCTION\ = \InstructionOne; \InstructionTwo; \DOT; \InstructionN\ be a sequence of instructions, and \PREDICATE\ be a path predicate  expressed in conjunctive form, capturing the branching conditions.
\begin{definition}
A symbolic execution state is represented by a tuple \STATE\ =
[\INSTRUCTION, \STACK, \PREDICATE].
A system configuration, \SYSTEM\ = \{\STATEONE, \STATETWO,
\DOT, \STATEN \}, is a set of  states.
\end{definition}
\subsection{Rules}
An instruction is defined by rules on a system configuration. There are two kinds of transitions: (1) \StateTrans\ internal transitions of a state, which are non-deterministic, and (2) \SystemTrans\ system transitions, which pick any state and replace it with all distinct states reachable by internal transitions, exploiting the non-determinism of the \StateTrans\ relation.

The semantics of the symbolic rules for the instructions adhere to the formal definitions of the Michelson language as outlined in \cite{michelson,michelson1}. In our model, these instructions are classified based on how we handle them into one-step, multi-step, blockchain, cryptographic, branch, and loop instructions.

One-step instructions directly modify the system state without branching, have a localized effect, and are modeled with a single rule.

Multi-step instructions involve the execution of a sub-sequence of instructions before returning to the main execution. The rule below models the \EXEC\ instruction, which executes a function specified as a sequence of instructions denoted by \INSTRUCTIONONE\ in the rule. This instruction applies the code of the function to the first element of the stack and then places the result back into the main execution.
\begin{mathpar}
  \inferrule[]
  { [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    \PREDICATE]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT \EMPTYSTACK, \PREDICATE']}{[(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE']}
\end{mathpar} 
Here, $\StateTrans^*$ represents the execution of multiple steps in a sequence of internal transitions.

Blockchain instructions, such as the \AMOUNT\
 instruction make implicit use of data from the transaction, the block and the current state of the blockchain. As these values are constant during the execution of a contract, we model them as fixed variables.

Branch instructions, such as \IF\{\INSTRUCTIONONE\}\{\INSTRUCTIONTWO\} consume a value at the top of the stack. If the branch condition
is True, it executes \INSTRUCTIONONE; otherwise, it executes
\INSTRUCTIONTWO. 
\begin{mathpar}
  \inferrule[]
  { }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    \{[\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne],   [\INSTRUCTIONTWO, \STACK, \PREDICATE\ \Wedge\ \NEG\
   \StackOne]\}} 
\end{mathpar}

The loop instructions, including \LOOP\ check the loop condition by examining the
top of the stack. These loop instructions may not terminate, just like while-loops. There are further loop instructions, like \ITER\ and
\MAP, specialized to traverse data structures like lists, sets, and maps. There are also
some instructions that implicitly require looping like \CONCAT, and
\SIZE. The latter kinds of loop instructions do always terminate.

Let us first take look at the terminate loops, such as the \ITER\ instruction, which iterates over a
list. The corresponding typing rule for the \ITER\ instruction is as follows: 
\begin{mathpar}
  \inferrule{\JTypeExpr\TEnv{\INSTRUCTION}{\TY \STACKCONCAT \TYA\ \SRightarrow\ \TYA}
  }{
      \JTypeExpr\TEnv{\ITER\ \INSTRUCTION}{\TYLIST\ \TY \STACKCONCAT \TYA\ \SRightarrow\ \TYA}
    }
\end{mathpar}
This rule ensures that if the loop body \INSTRUCTION\ has the type \TY\ \STACKCONCAT\ \TYA\ \SRightarrow\ \TYA, then the entire \ITER\ instruction, when applied to a stack with the top element of type \TY\ \TYLIST\ and the rest of the stack of type \TYA, produces a result stack of type \TYA.

The following rule handles the  \ITER\ loop for  the non-empty list\footnote{\ITER\ can also iterate over sets and maps, but the details are very similar to the list case and hence omitted.}, while the \ITER\ loop terminates when the list is empty.
\begin{mathpar}
  \inferrule[]
  { \HEAD, \STAIL\ \FRESH \\
    [\INSTRUCTIONONE,  (\HEAD, \TY) \STACKCONCAT\STACK, 
    \PREDICATE]
    \StateTrans^*
    [ \EMPTYSTACK,  \STACK', \PREDICATE']
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\
    \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \\
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\STAIL, \TYLIST\ \TY)
    \STACKCONCAT\STACK',  \PREDICATE' \Wedge  (\StackOne\
    \EQ\ \{\HEAD; \STAIL \}) ] 
  }
\end{mathpar}
It symbolically executes the instructions \INSTRUCTIONONE\ within the
\ITER\ block with the head of the list (\HEAD) replacing the list. The resulting state transitions to a
new stack (\STACK') and an updated predicate (\PREDICATE' $\Wedge$
(\StackOne\ \EQ\ \{\HEAD; \STAIL\})), where \STAIL\ represents the rest of the list.

If the list is concrete, the instruction terminates as it simply loops over the elements
of the list. How do we deal with \ITER\ if the list is a symbolic value? One solution could involve running the loop a certain number of times to obtain the symbolic value. However, we can defer that decision by abstracting the result.
